A Fundamental Measure Theory Functional for Hard-Sphere Contact Densities

### 2012-02-01 papers/contact/paper.tex
a bit of work on contact paper Tested everything.

@@

@@

We performed a Monte-Carlo simulation of the hard sphere fluid to
measure the contact density as a function of position for a few simple
geometries.  For each scenario, we compute and compare several
quantities.  We compare the Monte Carlo density with the DFT
prediction.  We also plot Monte Carlo results for the two contact
densities: the contact density evaluated at contact, and the contact
density evaluated at the sphere location.  Alongside these, we plot
the contact density of Fu and Wu~\cite{fu2005vapor-liquid-dft} (or
should it be Yu and
Wu\cite{yu2002fmt-dft-inhomogeneous-associating}?), the ``simple
contact density'' from Section~\ref{simple-contact}, and the ``contact
density at sphere'' from Section~\ref{contact-at-sphere}.  In
principle, the functional of Wu (FIXME) should match the
CenConDensity, while the DFT at sphere should match the ConDensity...

>>>>

% The following are papers that use a SAFT-based classical DFT with at
% least some of the terms purely local
\newcommand\saftlocaldft{felipe2001examination, gloor2002saft,%
  gloor2004accurate, clark2006developing, gloor2007prediction,%
  kahl2008modified, gross2009density}
% The following are papers that use a SAFT-based classical DFT with
% all the terms that should be non-local being non-local.
\newcommand\saftnonlocaldft{yu2002fmt-dft-inhomogeneous-associating,%
  fu2005vapor-liquid-dft,bryk2006density}

There has been considerable recent interest in using Statistical
Associating Fluid Theory (SAFT) to construct classical density
functionals to describe associating
fluids\cite{\saftlocaldft,\saftnonlocaldft}.  This approach has been
successful in qualitatively describing the dependence of surface
tension on temperature.  Unfortunately, most of these constructed
functionals\cite{\saftlocaldft} cannot be used to study arbitrary
inhomogeneous density distributions, because they use local
density approximations that are only valid for densities that vary
slowly over molecular length scales.

Constructing a classical density functional based on SAFT involves
rewriting each term in the free energy as a functional of an
inhomogeneous density.  The SAFT free energy consists of four terms:
\begin{align}
  A_\textit{SAFT} &= A_\textit{ideal} + A_\textit{HS} + A_\textit{chain} + A_\textit{disp} + A_\textit{assoc}
\end{align}
where $A_\textit{ideal}$ is the ideal-gas free energy, $A_\textit{HS}$
is the excess free energy of a hard-sphere fluid of monomers,
$A_\textit{chain}$ is the contribution to the free energy from the
chaining effect in a polymeric fluid, $A_\textit{disp}$ is the
dispersion contribution to the free energy, and $A_\textit{assoc}$ is
the free energy contribution from \emph{association}, which is to say,
hydrogen bonding.  Each of these terms has some dependence on the
inhomogeneity in the density.  The ideal gas is purely local, and in
fact must be the only purely local term in the free energy in order
for the functional to satisfy the contact-value theorem.  The
hard-sphere contribution contribution to the free energy has been
thoroughly studied, and is well approximated by the White Bear version
of the Fundamental-Measure Theory (FMT)
functional\cite{roth2002whitebear}.  A functional for the chain
contribution is has recently been introduced\cite{bryk2006density,
  gross2009density}, but is not needed for small associating molecules
such as water.  The dispersion term is long-range and thus has
significant dependence on the density distribution, but because its
relatively weak position dependence, is amenable to mean-field
approximations.  A functional for the association term of the free
energy was introduced by Wu and
Yu\cite{yu2002fmt-dft-inhomogeneous-associating}, and was used in
several subsequent functionals\cite{fu2005vapor-liquid-dft,
  bryk2006density}.  A key input (FIXME: term, factor?) to this
functional is the contact value of the correlation function of a
hard-sphere fluid.

The contact value of the correlation function---also known as the
cavity correlation function---is the the value of the
correlation function evaluated a distance equal to the hard-sphere
diameter.  Yu and Wu introduce an approximation based on the
fundamental measures of Fundamental Measure Theory (see
Section~\ref{sec:yuwu}).  In this paper, we will report on the
accuracy of this approximation by comparing with Monte Carlo
simulations of the inhomogeneous hard-sphere fluid.  In addition, we
will introduce and test differing approximations for the cavity
correlation function.

\textcolor{red}{Cavity correlation function!}

@@

\section{Gross' functional}\label{sec:gross}
We compare with the functional of Gross \emph{et
  al}\cite{gross2009density}, which is of the first kind, where the
contact density is evaluated at the sphere location.
\begin{align}
  n_\textit{contact} &= n_{0'} \frac{1 - \frac12 n_{3'}}{\left(1 - n_{3'}\right)^3}
\end{align}
where $n_{0'}$ and $n_{3'}$ are defined in a manner analogous to the
$n_0$ and $n_3$ of FMT, but with the hard-sphere radius replaced by
the hard-sphere diameter, so the integrals range over spheres that are
twice as large.

\section{Yu and Wu's functional}\label{sec:yuwu}

@@

We performed a Monte-Carlo simulation of the hard sphere fluid to
measure the contact density as a function of position for a few simple
geometries.  For each scenario, we compute and compare several
quantities.  We compare the Monte Carlo density with the DFT
prediction.  We also plot Monte Carlo results for the two contact
densities: the contact density evaluated at contact, and the contact
density evaluated at the sphere location.  Alongside these, we plot
the contact density of Fu and Wu~\cite{fu2005vapor-liquid-dft} (or
should it be Yu and Wu\cite{yu2002structures,
  yu2002fmt-dft-inhomogeneous-associating}?), the ``simple contact
density'' from Section~\ref{simple-contact}, and the ``contact density
at sphere'' from Section~\ref{contact-at-sphere}.  In principle, the
functional of Wu (FIXME) should match the CenConDensity, while the DFT
at sphere should match the ConDensity...
